---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

So my plan of action here is to look at my data using iNEXT. This package provides simple functions to compure and plot sample-size and coverage-based raretaction and extrapolation of species diversity individual based (abundance) data or sampling unit based (incident) data. 

```{r}
library(iNEXT)
```

To start, I realize that my data needs to be flipped, so that the data frame reads species as rows and site as columns. 

```{r}
#transpose data
tvisbun <- t(data_tvis)

#make dataframe from matrix
tvisbun <- as.data.frame(tvisbun)

```

Next, each individual site should be its own list. 

********DONT******* Run this - data must have the same number of rows, so I can't get rid of the 0s
```{r}
# Make data into one list
tvisbun <- as.list(tvisbun)

# Make each site into its own list, Sort in decreasing order, and remove zero values

# Avery Picnic Slope
APS <- (tvisbun$`Avery Picnic Slope`) 
  APS <- sort(APS, decreasing=TRUE) 
  APS <- APS[APS!=0]
  APS <- list(APS)

# Avery Picnic Upper
APU <- (tvisbun$`Avery Picnic Upper`)
  APU <- sort(APU, decreasing=TRUE) 
  APU <- APU[APU!=0]
  APU <- list(APU)
 
# Back of Baldy Bend   
BBB <- (tvisbun$`Back of Baldy Bend`) 
  BBB <- sort(BBB, decreasing=TRUE) 
  BBB <- BBB[BBB!=0]
  BBB <- list(BBB)

# Baldy Lunch
BL <- (tvisbun$`Baldy Lunch`) 
  BL <- sort(BL, decreasing=TRUE) 
  BL <- BL[BL!=0]
  BL <- list(BL)
 
# Bellview Bench 
BB <- (tvisbun$`Bellview Bench`) 
  BB <- sort(BB, decreasing=TRUE) 
  BB <- BB[BB!=0]
  BB <- list(BB)

#Bellview View
BV <- (tvisbun$`Bellview View`) 
  BV <- sort(BV, decreasing=TRUE) 
  BV <- BV[BV!=0]
  BV <- list(BV)

#billy's Hilly Lower
bHL <- (tvisbun$`billy's Hilly Lower`) 
  bHL <- sort(bHL, decreasing=TRUE) 
  bHL <- bHL[bHL!=0]
  bHL <- list(bHL)
  
#billy's Hilly Upper
bHU <- (tvisbun$`billy's Hilly Upper`) 
  bHU <- sort(bHU, decreasing=TRUE) 
  bHU <- bHU[bHU!=0]
  bHU <- list(bHU)

#Emerald Lake Lower
ELL <- (tvisbun$`Emerald Lake Lower`) 
  ELL <- sort(ELL, decreasing=TRUE) 
  ELL <- ELL[ELL!=0]
  ELL <- list(ELL)

#Emerald Lake Upper
ELU <- (tvisbun$`Emerald Lake Upper`) 
  ELU <- sort(ELU, decreasing=TRUE) 
  ELU <- ELU[ELU!=0]
  ELU <- list(ELU) 

#End of Gothic Hill 
EGH <- (tvisbun$`End of Gothic Hill`) 
  EGH <- sort(EGH, decreasing=TRUE) 
  EGH <- EGH[EGH!=0]
  EGH <- list(EGH)

#Kebler Pass
KP <- (tvisbun$`Kebler Pass`) 
  KP <- sort(KP, decreasing=TRUE) 
  KP <- KP[KP!=0]
  KP <- list(KP)
  
#Long Lake Parking Lot 1
LLP1 <- (tvisbun$`Long Lake Parking Lot 1`) 
  LLP1 <- sort(LLP1, decreasing=TRUE) 
  LLP1 <- LLP1[LLP1!=0]
  LLP1 <- list(LLP1)

#Long Lake Parking Lot 2
LLP2 <- (tvisbun$`Long Lake Parking Lot 2`) 
  LLP2 <- sort(LLP2, decreasing=TRUE) 
  LLP2 <- LLP2[LLP2!=0]
  LLP2 <- list(LLP2)
  
#Long Lake Parking Lot 3
LLP3 <- (tvisbun$`Long Lake Parking Lot 3`) 
  LLP3 <- sort(LLP3, decreasing=TRUE) 
  LLP3 <- LLP3[LLP3!=0]
  LLP3 <- list(LLP3)
  
#Long Lake Parking Lot 4
LLP4 <- (tvisbun$`Long Lake Parking Lot 4`) 
  LLP4 <- sort(LLP4, decreasing=TRUE) 
  LLP4 <- LLP4[LLP4!=0]
  LLP4 <- list(LLP4)
  
# Max Flied Meadow Lower
MFML <- (tvisbun$`Max Field Meadow Lower`) 
  MFML <- sort(MFML, decreasing=TRUE) 
  MFML <- MFML[MFML!=0]
  MFML <- list(MFML)

# Max Field Meadow Upper
MFMU <- (tvisbun$`Max Field Meadow Upper`) 
  MFMU <- sort(MFMU, decreasing=TRUE) 
  MFMU <- MFMU[MFMU!=0]
  MFMU <- list(MFMU)
  
# Ol' Dusty Road
ODR <- (tvisbun$`Ol' Dusty Road`) 
  ODR <- sort(ODR, decreasing=TRUE) 
  ODR <- ODR[ODR!=0]
  ODR <- list(ODR)
  
# Rosy Point Lower
RPL <- (tvisbun$`Rosy Point Lower`) 
  RPL <- sort(RPL, decreasing=TRUE) 
  RPL<- RPL[RPL!=0]
  RPL <- list(RPL)

# Rosy Point Upper
RPU <- (tvisbun$`Rosy Point Upper`) 
  RPU <- sort(RPU, decreasing=TRUE) 
  RPU <- RPU[RPU!=0]
  RPU <- list(RPU)

# Schofield/Vera Falls
SVF <- (tvisbun$`Schofield/Vera Falls`) 
  SVF <- sort(SVF, decreasing=TRUE) 
  SVF <- SVF[SVF!=0]
  SVF <- list(SVF)
  
# Slate River 2
SR <- (tvisbun$`Slate River 2`) 
  SR <- sort(SR, decreasing=TRUE) 
  SR <- SR[SR!=0]
  SR <- list(SR)

# Up on dat hill
Udh <- (tvisbun$`Up on dat hill`) 
  Udh <- sort(Udh, decreasing=TRUE) 
  Udh <- Udh[Udh!=0]
  Udh <- list(Udh)
  
# Vera Falls
VF <- (tvisbun$`Vera Falls`) 
  VF <- sort(VF, decreasing=TRUE) 
  VF <- VF[VF!=0]
  VF <- list(VF)
  
# Vera Falls 2
VF2 <- (tvisbun$`Vera Falls 2`) 
  VF2 <- sort(VF2, decreasing=TRUE) 
  VF2 <- VF2[VF2!=0]
  VF2 <- list(VF2)

```

RUN THIS INSTEAD - keeps zeros
```{r}
# Make data into one list
tvisbun <- as.list(tvisbun)

# Make each site into its own list, Sort in decreasing order, and remove zero values

# Avery Picnic Slope
APS <- (tvisbun$`Avery Picnic Slope`) 
  APS <- sort(APS, decreasing=TRUE) 
  APS <- list(APS)

# Avery Picnic Upper
APU <- (tvisbun$`Avery Picnic Upper`)
  APU <- sort(APU, decreasing=TRUE) 
  APU <- list(APU)
 
# Back of Baldy Bend   
BBB <- (tvisbun$`Back of Baldy Bend`) 
  BBB <- sort(BBB, decreasing=TRUE) 
  BBB <- list(BBB)

# Baldy Lunch
BL <- (tvisbun$`Baldy Lunch`) 
  BL <- sort(BL, decreasing=TRUE) 
  BL <- list(BL)
 
# Bellview Bench 
BB <- (tvisbun$`Bellview Bench`) 
  BB <- sort(BB, decreasing=TRUE) 
  BB <- list(BB)

#Bellview View
BV <- (tvisbun$`Bellview View`) 
  BV <- sort(BV, decreasing=TRUE) 
  BV <- list(BV)

#billy's Hilly Lower
bHL <- (tvisbun$`billy's Hilly Lower`) 
  bHL <- sort(bHL, decreasing=TRUE) 
  bHL <- list(bHL)
  
#billy's Hilly Upper
bHU <- (tvisbun$`billy's Hilly Upper`) 
  bHU <- sort(bHU, decreasing=TRUE) 
  bHU <- list(bHU)

#Emerald Lake Lower
ELL <- (tvisbun$`Emerald Lake Lower`) 
  ELL <- sort(ELL, decreasing=TRUE) 
  ELL <- list(ELL)

#Emerald Lake Upper
ELU <- (tvisbun$`Emerald Lake Upper`) 
  ELU <- sort(ELU, decreasing=TRUE) 
  ELU <- list(ELU) 

#End of Gothic Hill 
EGH <- (tvisbun$`End of Gothic Hill`) 
  EGH <- sort(EGH, decreasing=TRUE) 
  EGH <- list(EGH)

#Kebler Pass
KP <- (tvisbun$`Kebler Pass`) 
  KP <- sort(KP, decreasing=TRUE) 
  KP <- list(KP)
  
#Long Lake Parking Lot 1
LLP1 <- (tvisbun$`Long Lake Parking Lot 1`) 
  LLP1 <- sort(LLP1, decreasing=TRUE) 
  LLP1 <- list(LLP1)

#Long Lake Parking Lot 2
LLP2 <- (tvisbun$`Long Lake Parking Lot 2`) 
  LLP2 <- sort(LLP2, decreasing=TRUE) 
  LLP2 <- list(LLP2)
  
#Long Lake Parking Lot 3
LLP3 <- (tvisbun$`Long Lake Parking Lot 3`) 
  LLP3 <- sort(LLP3, decreasing=TRUE) 
  LLP3 <- list(LLP3)
  
#Long Lake Parking Lot 4
LLP4 <- (tvisbun$`Long Lake Parking Lot 4`) 
  LLP4 <- sort(LLP4, decreasing=TRUE) 
  LLP4 <- list(LLP4)
  
# Max Flied Meadow Lower
MFML <- (tvisbun$`Max Field Meadow Lower`) 
  MFML <- sort(MFML, decreasing=TRUE) 
  MFML <- list(MFML)

# Max Field Meadow Upper
MFMU <- (tvisbun$`Max Field Meadow Upper`) 
  MFMU <- sort(MFMU, decreasing=TRUE) 
  MFMU <- list(MFMU)
  
# Ol' Dusty Road
ODR <- (tvisbun$`Ol' Dusty Road`) 
  ODR <- sort(ODR, decreasing=TRUE) 
  ODR <- list(ODR)
  
# Rosy Point Lower
RPL <- (tvisbun$`Rosy Point Lower`) 
  RPL <- sort(RPL, decreasing=TRUE) 
  RPL <- list(RPL)

# Rosy Point Upper
RPU <- (tvisbun$`Rosy Point Upper`) 
  RPU <- sort(RPU, decreasing=TRUE) 
  RPU <- list(RPU)

# Schofield/Vera Falls
SVF <- (tvisbun$`Schofield/Vera Falls`) 
  SVF <- sort(SVF, decreasing=TRUE) 
  SVF <- list(SVF)
  
# Slate River 2
SR <- (tvisbun$`Slate River 2`) 
  SR <- sort(SR, decreasing=TRUE) 
  SR <- list(SR)

# Up on dat hill
Udh <- (tvisbun$`Up on dat hill`) 
  Udh <- sort(Udh, decreasing=TRUE) 
  Udh <- list(Udh)
  
# Vera Falls
VF <- (tvisbun$`Vera Falls`) 
  VF <- sort(VF, decreasing=TRUE) 
  VF <- list(VF)
  
# Vera Falls 2
VF2 <- (tvisbun$`Vera Falls 2`) 
  VF2 <- sort(VF2, decreasing=TRUE) 
  VF2 <- list(VF2)

```


Dang that took a while! Next is to bind these together into a list which I will use for iNEXT.

```{r}
# combine lists***
all <- as.list(c(APS, APU, BB, BBB, bHL, bHU, BL, BV, EGH, ELL, ELU, KP, LLP1, LLP2, LLP3, LLP4, MFML, MFMU, ODR, RPL, RPU, SR, SVF, Udh, VF, VF2))

# remove sites with only 1 species
two <- as.list(c(APS, APU, BB, BBB, bHL, BL, BV, EGH, ELL, ELU, KP, LLP1, LLP2, LLP3, LLP4, MFML, MFMU, ODR, RPL, RPU, SR, SVF, Udh, VF2))

# make into dataframe
two <- as.data.frame(two)
```


Now I am going to create the abundance graph.
*I'm going to use q=1 (Shannon's) for simplicity to begin but will go back later to determine what I should actually use.

```{r}
out <- iNEXT(two, q=1, datatype="abundance", endpoint=NULL)

# Sample-size-based R/E curves, separating by "order"
ggiNEXT(out, type=1, facet.var="none") + 
  xlim(0,100)
```





One of the analyses that I need to run will use the total visitation by species comparing each year. For now I have the 2012 and 2018 data.

First, I need to get the totals from each year:

```{r}
# make dataframe for 2012
tvisbun2 <- t(data_tvis) #transpose data
tvisbun2 <- as.data.frame(tvisbun2)
tvisbun2 <- tvisbun2[-c(3),]

# sum row totals in new dataframe
totals12 <- as.data.frame(rowSums(tvisbun2))

#sort in decreasing order
totals12 <- sort(totals12$`rowSums(tvisbun2)`, decreasing=TRUE)

# make into list
#totals12 <- list(c(totals12))

# make dataframe for 2018
tvisbun8 <- t(data_tvis8) #transpose data
tvisbun8 <- as.data.frame(tvisbun8)

## sum row totals in new dataframe
totals18 <- as.data.frame(rowSums(tvisbun8))

#sort in decreasing order
totals18 <- sort(totals18$`rowSums(tvisbun8)`, decreasing=TRUE)

# make into list
#totals18 <- list(c(totals18))

```

Now I'll merge them into one dataframe

```{r}
both <- cbind(totals12, totals18)
```


```{r}
out2 <- iNEXT(both, q=c(0, 1, 2), datatype="abundance", endpoint=NULL)

# Sample-size-based R/E curves, separating by "order""
ggiNEXT(out2, type=1, facet.var="site") + 
  xlim(0,175)
```


```{r}
# Sample-size-based R/E curves, separating by "order""
ggiNEXT(out2, type=1, facet.var="order") + 
  xlim(0,175)
       
```























For the sample-unit based incidence data, the first row should be column totals. 

```{r}
#create dataframe with row totals
totals <- as.data.frame(colSums(tvisbun))
totals <- t(totals)

#bind the totals dataframe with the original dataframe to create incidence data
tvisinc <- rbind(totals, tvisbun)

```



How about incidence graph?

```{r}
out2 <- iNEXT(tvisinc, q=1, datatype="incidence_freq", endpoint=NULL)

ggiNEXT(out2, type=1, facet.var="none")
```







